---
title: "Assignment 01"
author: "Karisa Ke"
format: html
embed-resources: true
editor: visual
---

```{r}
library(data.table)
data_2002 <- fread("/Users/yunqingke/Downloads/PM2.5_2002.csv")
data_2022 <- fread("/Users/yunqingke/Downloads/PM2.5_2022.csv")
```

```{r}
dim(data_2002)
```

```{r}
head(data_2002)
```

```{r}
tail(data_2002)
```

```{r}
str(data_2002)
```

```{r}
dim(data_2022)
```

```{r}
head(data_2022)
```

```{r}
tail(data_2022)
```

```{r}
str(data_2022)
```

```{r}
summary(data_2002$'Daily Mean PM2.5 Concentration')
```

```{r}
summary(data_2022$'Daily Mean PM2.5 Concentration')
```

Summary: There are 15976 rows and 20 columns in the data_2002 dataset. There are 56140 rows and 20 columns in the data_2022 dataset. The variable names and types are the same for both datasets.

```{r}
combined_data <- rbind(data_2002, data_2022)
```

```{r}
dim(combined_data)
```

```{r}
combined_data$year <- as.integer(substr(combined_data$Date, 7, 10))

```

```{r}
dim(combined_data)
```

```{r}
head(combined_data)
```

```{r}
tail(combined_data)
```

```{r}
names(combined_data)[names(combined_data) == "Daily Mean PM2.5 Concentration"] <- "PM2.5"
```

```{r}
head(combined_data)
```

```{r}
library(leaflet)
```

```{r}
# Create a color column based on the year
combined_data$marker_color <- ifelse(combined_data$year == 2002, "blue", "red")

# Create a basic map with markers and a legend
map <- leaflet(combined_data) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~SITE_LONGITUDE,
    lat = ~SITE_LATITUDE,
    color = ~marker_color,
    radius = 4,
    label = ~year
) %>%
  addLegend(
    values = c("blue", "red"),
    colors = c("blue", "red"),
    labels = c("2002", "2022"),
    title = "Year"
)

# Print the map
map
```

The monitoring sites are mainly located by the coastline.

```{r}
# Check for missing values
missing_values <- sum(is.na(combined_data$PM2.5))
```

```{r}
missing_values
```

```{r}
# Check for implausible values (e.g., negative values)
implausible_values <- sum(combined_data$PM2.5 < 0)
```

```{r}
implausible_values
```

```{r}
# Calculate proportions
total_records <- nrow(combined_data)
missing_proportion <- missing_values / total_records
implausible_proportion <- implausible_values / total_records
```

```{r}
missing_proportion
```

```{r}
implausible_proportion
```

```{r}
cat("Missing PM2.5 values:", missing_proportion * 100, "%\n")
cat("Implausible PM2.5 values:", implausible_proportion * 100, "%\n")
```

```{r}
# at state level
state_summary <- combined_data[, .(Average_PM2.5 = mean(PM2.5, na.rm = TRUE)), by = .(year)]

# Line plot of PM2.5 concentrations over the years
library(ggplot2)
ggplot(state_summary, aes(x = year, y = Average_PM2.5, group = 1)) +
  geom_line() +
  labs(x = "Year", y = "Average PM2.5 Concentration (µg/m³)", title = "Average PM2.5 Concentrations in California (2002-2022)")
```

The line plot shows the average PM2.5 concentrations in California from 2002 to 2022, showing that the average PM2.5 level was decreasing over the time.

```{r}
# at county level
library(ggplot2)
ggplot(combined_data, aes(x = combined_data$COUNTY, y = PM2.5, fill = factor(year))) +
  geom_boxplot() +
  labs(x = "County", y = "PM2.5 Concentration (µg/m³)", title = "PM2.5 Concentrations by County (2002 vs. 2022)") +
  scale_fill_manual(values = c("blue", "red")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The boxplot compares PM2.5 concentrations across different counties in California for the years 2002 and 2022. There are several counties that have extremely high PM2.5 levels.

```{r}
# site in LA
la_data <- combined_data[COUNTY == "Los Angeles"]

# Histogram of PM2.5 concentrations at specific sites in Los Angeles for 2002 and 2022
ggplot(la_data, aes(x = PM2.5, fill = factor(year))) +
  geom_histogram(binwidth = 5, position = "dodge") +
  labs(x = "PM2.5 Concentration (µg/m³)", y = "Frequency", title = "PM2.5 Concentrations at Specific Sites in Los Angeles (2002 vs. 2022)") +
  scale_fill_manual(values = c("blue", "red"))
```

The histograms show the distribution of PM2.5 concentrations at specific monitoring sites in Los Angeles for the years 2002 and 2022. We can see that overall the year of 2022 has much high PM2.5 levels than the year of 2002.
